---
title: "Ometa_Organic_Amendment_Meta_Analysis"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

#### Goal: Analalyze data on soil organic amendment impact on soil carbon pools > 20cm in depth

#### Notes:
       #echo=false supresses code; include=false supressed output
        #Last working on:
        1) COMPLETE Make categories for each variable = cropping system (monocrop, rotation), crop type (grain, row, legume, forage), years of application, type of amendment, soil pH, SOC content intial, clay content, latitute, C rate applied, dry weight, N rate 
        2) Separating out POM/MAOM data

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


```{r, prelimary - load packages, include=FALSE}
#include=FALSE supress this output when knitting doc
uzdir<-("/Users/macuser/Dropbox/9Meta_analysis_organic/Statistical_meta_analysis") #save long name as user directory uzdir
setwd(paste(uzdir))  #set your working directory/file
source(file="libraries_20190816.R") #install packages, use old vegan 
```

```{r, climate data, eval=FALSE}
#http://www.worldclim.org/formats1

#tmean , prec ; 12 data layers 1 for each month
#BIO1 = Annual Mean Temperature; BIO2 = Mean Diurnal Range (Mean of monthly (max temp – min temp)); BIO3 = Isothermality (BIO2/BIO7) (* 100)
#BIO4 = Temperature Seasonality (standard deviation *100); BIO5 = Max Temperature of Warmest Month, BIO6 = Min Temperature of Coldest Month
#BIO7 = Temperature Annual Range (BIO5-BIO6), BIO8 = Mean Temperature of Wettest Quarter, BIO9 = Mean Temperature of Driest Quarter
#BIO10 = Mean Temperature of Warmest Quarter, BIO11 = Mean Temperature of Coldest Quarter, BIO12 = Annual Precipitation
#BIO13 = Precipitation of Wettest Month, BIO14 = Precipitation of Driest Month, BIO15 = Precipitation Seasonality (Coefficient of Variation)
#BIO16 = Precipitation of Wettest Quarter, BIO17 = Precipitation of Driest Quarter, BIO18 = Precipitation of Warmest Quarter, BIO19 = Precipitation of Coldest Quarter

#download resolution of climate data desired: http://worldclim.org/version2 #version 2 = 1970-2000
  #or version 1.4, under generic grid format, downlaod bioclim 2.5 from https://www.worldclim.org/current

#r <- raster::getData("worldclim",var="bio",res=2.5)


# also possible to get future climate data: 
# https://rdrr.io/cran/raster/man/getData.html

#r <- r[[c(1,12)]] #Bio1 and Bio12 selected
#names(r) <- c("Temp","Prec")

#lats <- c(df1$lat.dec.degree) #use spTransform if not in WGS 84 lat/lon (EPSG 4326)
#longs <- c(df1$long.dec.degree) 
# coords <- data.frame(x=longs,y=lats)

# points <- SpatialPoints(coords, proj4string = r@crs)
# values <- raster::extract(r,points)

#df.clim <- cbind.data.frame(coordinates(points),values); head(df.clim,2)
#df.clim$Temp<-df.clim$Temp/10 #WorldCLim data hs a scale factor of 10 for temp
#WriteXLS(df.clim, "ometa_climate.xls") #specific file here

```

```{r, load data and subset, include=FALSE}
#echo=FALSE do not incldue code when knitting doc
#include=FALSE supress this output when knitting doc

#rm(list = ls()) # if you want to reset your environment by deleating everything
df1<-read.csv("Ometa_table.csv", header=T, na.strings= c(""," ", "na", "nd", "NA"), skip=2)
str(df1)
ncol(df1)
nrow(df1)
#colnames(df1)

#number of studies included
length(unique(df1$ACC))
length(unique(df1$country))
length(unique(df1$country))
df1$X_a<-as.numeric(df1$X_a); df1$X_cf<-as.numeric(df1$X_cf); df1$X_cmin<-as.numeric(df1$X_cmin)
```

```{r, initial plots, echo=FALSE}
plot(df1$country); axis(2,cex.axis=1)
plot(df1$depth.end)
plot(df1$param_group, main="Parameters - Reponse Variables")
plot(df1$tillage_type, main = "Tillage Groups")
plot(df1$crop_group, main= "Crop Types")
plot(df1$crop_system)
hist(df1$lat.dec.degree, main= "Latitude")
     hist(df1$elevation, main= "Elevation", xlab="Elevation (m)")
     plot(df1$MAP, df1$MAT)
     hist(df1$year.zero, main="First Year of Study", xlab="Year")
     hist(df1$MAP, main="Precipitation", xlab="MAP (mm)")
      hist(df1$MAT, main="Temperature", xlab="MAT (degrees C)")
hist(df1$clay, main="Percent Clay", xlab="Clay (%)")
hist(df1$depth.end)
plot(df1$crop_system)
plot(df1$duration_group, main="Study Duration", xlab= "Years")
    plot(df1$irrigation_group, main ="Cateogry of Irrigation")
   plot(df1$till_depth_group, main ="Tillage depths")
   hist(df1$pH, xlab="pH")  
   hist(df1$SOC_g.kg, main="Initial SOC", xlab="SOC g/kg")

   
#levels(df1$param)
#levels(df1$crop)
#names(df1)

#columsn of importance

df1 %>% dplyr::select(ACC, crop_group, crop_system, depth.start ,  depth.end ,  param ,  param_group ,  units , amend ,  irrigation ,  tillage_type , tillage_depth, till_depth_group, country , elevation ,  lat_group ,  MAT ,  MAP ,  duration_group,  duration_even ,  clay_group ,  pH_group ,  SOC_group, X_cf, SE_cf, SD_cf, N_cf, X_a, SE_a, SD_a, N_a, X_cmin, SE_cmin, SD_cmin, N_cmin ) -> df
  
df1 %>% dplyr::select(X_cf, SE_cf, SD_cf, N_cf, X_a, SE_a, SD_a, N_a, X_cmin, SE_cmin, SD_cmin, N_cmin)-> df.result
```

```{r, preliminary response ratios, eval=FALSE}
#preliminary ratios for effect size
hist(log(df$X_a))

ratio_min<-log(df1$X_a/df1$X_cmin)
    percent.change.min<-(exp(ratio_min) -1)*100
    mean(percent.change.min, na.rm=TRUE)

ratio_fert<-log(df1$X_a/df1$X_cf)
    percent.change.fert<-(exp(ratio_fert) -1)*100
mean(percent.change.fert, na.rm=TRUE)

df2<-cbind(df1, ratio_min, ratio_fert)

df.c<-subset(df2, param_group == "C")
 
plot()
```

```{r, determine ranges, eval=FALSE}
df<-df2

#crop
levels(df$crop)

#crop_system

#depth.end
#amend
#irrigaiton
#country
#lat.long
#eleveation
#MAP_av
#MAT_av
#soil.texture
#clay
#pH
#SOC_int
#C.rate


```


#### Previous analyses
Additional code is also saved within the .Rmd script from previous statistical analyses for easy access.
```{r, eval=FALSE, include=FALSE}
names(df1) <- gsub(".Concentration", ".pXRF", names(df1))
names(df1) <- gsub(".Error1s", ".E", names(df1))
df1$layer_bot<-as.factor(df1$layer_bot); df1$veg_age<-as.factor(df1$veg_age); df1$sample_position<-as.factor(df1$sample_position)
#examine imported data #colnames(df1) #str(df1) #tail(df1)

#downsize dataframe
df.factor<-df1[, c(1:24)]  #pick variables of interest names(df.factor)
df.data<-df1[, c(76:length(df1))] #choose columns with data variables of interest
df.chem<-df1[,76:82]
df.xrf<- as.data.frame(df.data %>% dplyr::select(contains('.pXRF')))
df.xrf<- mutate_all(df.xrf, function(x) as.numeric(as.character(x))); #str(df.xrf)
names(df.xrf)
df.error1<-df.data %>% dplyr::select(contains('.E')) # str(df.error1)
df.error<-cbind(df.factor, df.chem, df.error1)
df.percent<-cbind(df.factor,df.chem, df.xrf) #names(df.percent)
df.wide<-cbind(df.factor, df.chem, df.data)

df.percent$layer_bot<-as.integer(df.percent$layer_bot)

df.v<-subset(df.percent, veg=="V_vineyard" & sample_method=="A_auger_10.16cm" & site_name== "2CIEPA") #subset of vineyards only
names(df.v)
df.vine<-df.v[rowSums(is.na(df.v[ , 32:length(df.v)])) == 0, ] #remove rows with all NAs
#str(df.vine)
df.v.error<-subset(df.error, veg=="V_vineyard" & sample_method=="A_auger_10.16cm" & site_name== "2CIEPA" ) #subset of vineyards only
df.vine.error<-df.v.error[rowSums(is.na(df.v.error[ , 32:length(df.v.error)])) == 0, ]
#put into long format names(df.vine)
df.long.ppm<-gather(data=df.vine, key=Element, value=PPM, c(21:length(df.vine)))
df.long.error<-gather(data=df.vine.error, key=Element, value=SD, c(21:length(df.vine.error)))
#str(df.long.error); str(df.long.ppm)
names(df.vine)
SD<-as.numeric(df.long.error$SD)

df.long<-cbind(df.long.ppm, SD)
  #colnames(df.long); str(df.long)

#after data wrangling is finished, save dataframe as simply df
  df<-df.long #str(df)
  
#make ppm and error Numeric
 df$PPM<-as.numeric(df$PPM)
```

```{r, check for normality, eval=FALSE, include=FALSE}
str(df.vine)
#colnames(df.vine)

#check C and N data
#histogram
hist(df.vine$C)
hist(df.vine$N)
plot(df.vine$layer_bot, df.vine$C)
plot(df.vine$sample_position, df.vine$N)

#qqplot
qqnorm(df.vine$C); qqline(df.vine$C)
qqnorm(df.vine$N); qqline(df.vine$N)

#shapiro.wilk test
shapiro.test(log(df.vine$C)) #log transform C
shapiro.test(sqrt(df.vine$N)) #sqrt transform N

sC<-log(df.vine$C)
sN<-sqrt(df.vine$N)

mean(df.vine$C, na.rm=T)
mean(df.vine$N, na.rm=T)

### Check normality of XRF data and pH
#plot
plot(df.vine$pH_CaCl2, df.vine$pH_H2O)

#histogram
hist(df.vine$pH_CaCl2)
#hist(df.vine$electrical_conductivity)

#plot
plot(df.vine$layer_bot, df.vine$pH_CaCl2)
plot(df.vine$sample_position, df.vine$pH_CaCl2)

#qqplot
qqnorm(df.vine$pH_CaCl2); qqline(df.vine$pH_CaCl2)

#shapiro.wilk test
shapiro.test(df.vine$pH_CaCl2)

#plot
mean(df.vine$pH_CaCl2, na.rm=T)
mean(df.vine$pH_H2O, na.rm=T)
plot(df.vine$pH_CaCl2, df.vine$pH_H2O)

#remove columns with zeros
df.sub<-df.vine[28:length(df.vine)][, colSums(df.vine[28:length(df.vine)] != 0, na.rm=T) > 0]
#run test for normality
df.normal.metals <- as.data.frame(sapply(df.sub, function(x) shapiro.test(log(x+1)))); 
df.norm.all<-as.data.frame(t(df.normal.metals))[1:2] #transpose and take only stat and p-value
df.normal<-df.norm.all[df.norm.all$p.value > .05,] #check which metals are NORMAL p value > .05
df.normal #normal variables = Fe, K, Rb, Ti, V
#log x+1 normal = Co, Fe, K, Mn, Rb, Ti, V, Zr

#look at histograms for metals
sapply(df.sub, function(x) hist(x))

colnames(df.sub) #30 metals
m<-df.sub$Zr.pXRF ; hist(m) 
             #all zeros: Ag, Bi, Cd, Sb, Sn 
             #log: Co, Cu p= .03, Mn, Rb, Zr 
             #one outlier: Pb, Sn (only 1 value=almost all zeros); Zn 2 outliers
             #normal variables = Fe, K, Rb, Ti, V
             #Other: Al, As, Ca, Cr, Hg, LE, Ni, P, S, Sr, Th, W, Y, Mg
            #Al several outliers ; As 5 outliers, patterns?, Ca 6 outliers, Cr many outliers, Hg few  values registered; LE not normal, curving qqplot; Mg four outliers; Mo very low values <17, same values repeated ; Nb same values repeated; Ni long tail; P set of outliers; S set of outliers; Sr set of oultiers; Th 7 outliers all about under 20; W very few values <25%; Y man of same values <18; 

#log(Cu), boxcox/KW(Pb), boxcox/KW(Ni), Zn, boxcox/KW(As), outliers-Mg, boxcox(Cr), Fe, log(Mn) (NO Cd values)

qqnorm(m); qqline(m)
lm<-log(m); shapiro.test(lm)
qqnorm(lm); qqline(lm)
sm<-sqrt(m); shapiro.test(sm)
```

```{r, C linear regression models / ANOVA, eval=FALSE, echo=FALSE}
 #set up liner mixed effects model, with Block as a fixed effect
#lmer.c<-lmer (log(C) ~ sample_position*layer_bot+ (1|veg_age) 
#                    + (1|veg_age:sample_position), data=df.vine)
#  resid.c<- resid(lmer.c) #plot residual
#  plot(resid.c);abline(0,0) #look normal, evenly distributed
# anova(lmer.c, ddf="Kenward-Roger") #run analysis of variance, ddf = posthoc correction for mixed model

#test without vineyard age as fixed effect
lm.c<-lm (log(C) ~ veg_age*sample_position*layer_bot, data=df.vine)
  resid.clm<- resid(lm.c) #plot residual
  plot(resid.clm);abline(0,0) #look normal 
anova(lm.c) #run analysis of variance, ddf = posthoc correction for mixed model

 #model selection
# Fit the full model 
full.model.c <- lm(sqrt(C) ~ veg_age*layer_bot*sample_position, data=df.vine)
# Stepwise regression model
step.model.c <- stepAIC(full.model.c, direction = "both", trace = FALSE)
summary(step.model.c)

#pairwise contrasts lmer
cont3<-lsmeans(lmer.c, pairwise ~ sample_position*layer_bot);
cont3.table<-summary(cont3$contrasts); cont3.table
cont3.table[(cont3.table$p.value<=.05),]

#pairwise contrasts lm (vine age as a random treatment effect)
cont1<-lsmeans(lm.c, pairwise ~ sample_position*layer_bot|veg_age);
cont1.table<-summary(cont1$contrasts); cont1.table
cont1.table[(cont1.table$p.value<=.05),]

cont2<-lsmeans(lm.c, pairwise ~ veg_age*layer_bot|sample_position);
cont2.table<-summary(cont2$contrasts); cont2.table
cont2.table[(cont2.table$p.value<=.05),]
```

```{r, N linear regression models / ANOVA, eval=FALSE, echo=FALSE}
 #set up liner mixed effects model, with Block as a fixed effect
#lmer.n<-lmer (sqrt(N) ~ sample_position*layer_bot+ (1|veg_age) 
#                    + (1|veg_age:sample_position), data=df.vine)
#  resid.n<- resid(lmer.n) #plot residual
#  plot(resid.n);abline(0,0) #look normal, evenly distributed
#anova(lmer.n, ddf="Kenward-Roger") #run analysis of variance, ddf = posthoc correction for mixed model

#test without vineyard age as a fixed effect
lm.n<-lm (sqrt(N) ~ veg_age*sample_position*layer_bot, data=df.vine)
  resid.nlm<- resid(lm.n) #plot residual
  plot(resid.nlm);abline(0,0) #look normal 
anova(lm.n) #run analysis of variance, ddf = posthoc correction for mixed model

 #model selection
# Fit the full model 
full.model.n <- lm(sqrt(N) ~ veg_age*layer_bot*sample_position, data=df.vine)
# Stepwise regression model
step.model.n <- stepAIC(full.model.n, direction = "both", trace = FALSE)
summary(step.model.n)

#pairwise contrasts lmer
cont3<-lsmeans(lmer.n, pairwise ~ sample_position*layer_bot);
cont3.table<-summary(cont3$contrasts); cont3.table
cont3.table[(cont3.table$p.value<=.05),]

#pairwise contrasts lm (vine age as a random treatment effect)
cont1<-lsmeans(lm.n, pairwise ~ sample_position*layer_bot|veg_age);
cont1.table<-summary(cont1$contrasts); cont1.table
cont1.table[(cont1.table$p.value<=.05),]

cont2<-lsmeans(lm.n, pairwise ~ veg_age*layer_bot|sample_position);
cont2.table<-summary(cont2$contrasts); cont2.table
cont2.table[(cont2.table$p.value<=.05),]
```

```{r, C and N graph, eval=FALSE, echo=FALSE}

df.c.plot<-df.vine
names(df.c.plot)[c(11,22, 24)]<-c("Months",  "Position", "Depth")

#graph mean sample position, depth, months
mean.c <- aggregate(C ~ Months + Position + Depth, df.plot, function(x) c(mean = mean(x)), na.action=na.omit);str(mean.c)
se.c<- aggregate(C~ Months + Position + Depth, df.c.plot, function(x) c(se = sd(x)/sqrt(length(x))), na.action=na.omit);se.c;
tab.c<-cbind(mean.c, se.c$C);names(tab.c)[1:5]<-c("Months",  "Position", "Depth", "C", "SE")
str(tab.c)

#carbon
p.c<-ggplot(data=df.plot, aes(x=Months, y=C, fill=Position)) +
    # geom_hline(yintercept=100, linetype=1, color="darkgray", size=1)+
    #geom_errorbar(aes(ymin=C-SE, ymax=C+SE), width=.1)+
  geom_boxplot(size=1, aes(fill = Position),color="black")+ #, shape=21)+
 facet_wrap(Depth~.,  ncol=2) #scale="free"
p.c

p.c.pos<-ggplot(data=df.plot, aes(x=Months, y=C, fill=Depth)) +
    # geom_hline(yintercept=100, linetype=1, color="darkgray", size=1)+
    #geom_errorbar(aes(ymin=C-SE, ymax=C+SE), width=.1)+
  geom_boxplot(size=1, aes(fill = Depth),color="black")+ #, shape=21)+
 facet_wrap(Position~.,  ncol=2) #scale="free"
p.c.pos

#nitrogen
p.n<-ggplot(data=df.plot, aes(x=Months, y=N, fill=Position)) +
    # geom_hline(yintercept=100, linetype=1, color="darkgray", size=1)+
    #geom_errorbar(aes(ymin=C-SE, ymax=C+SE), width=.1)+
  geom_boxplot(size=1, aes(fill = Position),color="black")+ #, shape=21)+
 facet_wrap(Depth~.,  ncol=2) #scale="free"
p.n

p.n.pos<-ggplot(data=df.plot, aes(x=Months, y=C, fill=Depth)) +
    # geom_hline(yintercept=100, linetype=1, color="darkgray", size=1)+
    #geom_errorbar(aes(ymin=C-SE, ymax=C+SE), width=.1)+
  geom_boxplot(size=1, aes(fill = Depth),color="black")+ #, shape=21)+
 facet_wrap(Position~.,  ncol=2) #scale="free"
p.n.pos


```

```{r, Metals - check for correlations between response variables, eval=FALSE, echo=FALSE}
#Pearson correlation for normal variables
df.corr<-cbind(df.vine[26:length(df.vine)]) 
spear<-rcorr(x=as.matrix(df.corr), y=as.matrix(df.corr), type = "spearman") #type can be pearson/spearman 
  tab.corr<-as.data.frame(spear$r)[1:30, 1:30] #cut repeated set of variables
  rownames(tab.corr); colnames(tab.corr)
  tab.p<-as.data.frame(spear$P)[1:30, 1:30] #cut repeated set of variables
  corr.table<-rbind(tab.corr, tab.p) 
  corr.order<- corr.table[order(row.names(corr.table)), ]
  write.table(corr.order, "nexus_xrf_spearman.csv", sep =",") #rows with .1 are the p-values
#str(corr.df) 
  
#>.75	Zn/Cu ; Ti/Fe
#>.60			Rb/K, Si/Al,	Ca/Cu, Co/Fe,	Mn/Co, Cu/S, Mn/Fe,LE/S,Th/Mo,Nb/Y
```

```{r, METALS check RCRA metals/standards, eval=FALSE, echo=FALSE}
#   Element	D-Limit (ppm)	EPA alllowable limits
#24	Cr    	<10ppm	      5x
#33	As	    <5ppm	        5x  6ppm  .1-40
#34	Se	    <5ppm	        1
#47	Ag	    <10ppm	      5
#48	Cd	    <10ppm	      1   .35 .01-2
#56	Ba	    <20ppm	      100x
#80	Hg	    <5ppm	        0.2x
#82	Pb	    <5ppm	        5x

#Percent of smaples over RCRA limit for 'hazardous' waste
Cr.RCRA<- df.vine$Cr.pXRF > 5; length(Cr.RCRA[Cr.RCRA==TRUE])/length(Cr.RCRA)*100 # = 100% 
As.RCRA<- df.vine$As.pXRF > 5; length(As.RCRA[As.RCRA==TRUE])/length(As.RCRA)*100 # = 92%
Se.RCRA<- df.vine$Se.pXRF > 5; length(Se.RCRA[Se.RCRA==TRUE])/length(Se.RCRA)*100 # = 6.5% 
#df.vine$Ag.pXRF # not detected
#df.vine$Cd.pXRF #not detected
Hg.RCRA<- df.vine$Hg.pXRF > .2; length(Hg.RCRA[Hg.RCRA==TRUE])/length(Hg.RCRA)*100 # = 32% 
Pb.RCRA<- df.vine$Pb.pXRF > 5; length(Pb.RCRA[Pb.RCRA==TRUE])/length(Pb.RCRA)*100 # = 100%
colnames(df.vine)
 
df.vine.error$Pb.E
df.vine.error$Mg.E

#Cu, Pb, Ni, Zn, Cd
#As, Mg, Cr, Fe, Mn
```

```{r, Anova - seperate for each metals of interest, eval=FALSE, echo=FALSE}
#log(Cu), boxcox/KW(Pb), boxcox/KW(Ni), Zn, boxcox/KW(As), outliers-Mg, boxcox(Cr), Fe, log(Mn) 
#(NO Cd values)

#ANOVA for each metal

As.model <-(aov(As~layer_bot*sample_position*veg_age, df.m)) 
summary(As.model)

Cr.model <-(aov(Cr~layer_bot*sample_position*veg_age, df.m)) 
summary(Cr.model)
#pairwise comparisions of means
cont1<-lsmeans(Cr.model, pairwise ~ layer_bot*veg_age*sample_position);
cont1.table<-summary(cont1$contrasts); cont1
cont2.table[(cont2.table$p.value<=.05),]


Cu.model <-(aov(Cu.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Cu.model)

Fe.model <-(aov(Fe.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Fe.model)

Mg.model <-(aov(Mg.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Mg.model)

Mn.model <-(aov(Mn.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Mn.model)

Ni.model <-(aov(Ni.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Ni.model)

Pb.model <-(aov(Pb.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Pb.model)

Zn.model <-(aov(Zn.pXRF~layer_bot*sample_position*veg_age, df.m)) 
summary(Zn.model)
```

```{r, MANOVA of metals, eval=FALSE, echo=FALSE}
# variable to combine: REMOE: Cd.pXRF, Bi.pXRF,  Ag.pXRF, Sb.pXRF, Sn.pXRF,
df.m<-dplyr::select(df.vine, veg_age, layer_bot, sample_position, Al.pXRF, As.pXRF, Ca.pXRF, Co.pXRF, Cr.pXRF, Cu.pXRF, Fe.pXRF, Hg.pXRF, K.pXRF, LE.pXRF, Mg.pXRF, Mn.pXRF, Mo.pXRF, Nb.pXRF,Ni.pXRF, P.pXRF, Pb.pXRF, Rb.pXRF, S.pXRF,  Se.pXRF, Si.pXRF,  Sr.pXRF, Th.pXRF, Ti.pXRF, U.pXRF, V.pXRF, W.pXRF, Y.pXRF, Zn.pXRF, Zr.pXRF)
names(df.m)<-str_replace(names(df.m), ".pXRF", "")

#Cu, Pb, Ni, Zn,  As, Mg, Cr, Fe, Mn (NO Cd values)
m1<-dplyr::select(df.vine, As.pXRF,   Cr.pXRF, Cu.pXRF, Fe.pXRF,  Mg.pXRF, Mn.pXRF,  Ni.pXRF, Pb.pXRF,  Zn.pXRF)
names(m1)<-str_replace(names(m1), ".pXRF", "")
names(df.vine)
m2<-dplyr::select(df.vine, veg_age, layer_bot, sample_position, Al.pXRF,Ca.pXRF, Co.pXRF, Hg.pXRF, K.pXRF, LE.pXRF, Mo.pXRF, Nb.pXRF, P.pXRF, Rb.pXRF, S.pXRF,  Se.pXRF, Si.pXRF,  Sr.pXRF, Th.pXRF, Ti.pXRF, U.pXRF, V.pXRF, W.pXRF, Y.pXRF,Zr.pXRF, C, N, pH_CaCl2, electrical_conductivity)
names(m2)<-str_replace(names(m2), ".pXRF", "")
  #df.vine$Zn.pXRF #Zn average in soil 90ppm - not interesting
  #df.vine$Mn.pXRF #Mn also under sol average of 1000ppm

df.m$veg_age<-as.factor(df.m$veg_age)
df.m$layer_bot<-as.factor(df.m$layer_bot)

### Permanova 
   #not working, with metals of interest as the reponse, and others as the data
   #Correlated varibles?
#>.75	Zn/Cu ; Ti/Fe
#>.60			Rb/K, Si/Al,	Ca/Cu, Co/Fe,	Mn/Co, Cu/S, Mn/Fe,LE/S,Th/Mo,Nb/Y

names(m1)
names(m2)

```

```{r, graph metals, eval=FALSE, echo=FALSE}
pd <- position_dodge(.5)

#prep dataframe for the metals graph
df.vine.long<-subset(df.long, veg=="V_vineyard" & sample_method=="A_auger_10.16cm" & site_name== "2CIEPA") #subset of vineyards only
df.vine.long$Element<-as.factor(df.vine.long$Element)
names(df.vine.long)
str(df.vine.long)

#choose elements of interest
df.plot<-subset(df.vine.long, Element==c("As.pXRF","Cr.pXRF","Cu.pXRF" ,"Ni.pXRF", "Pb.pXRF", "Mn.pXRF", "Zn.pXRF", "Mg.pXRF", "Fe.pXRF"));str(df.plot)
#df.plot$layer_bot<-as.factor(df.plot$layer_bot); 
df.plot$veg_age<-as.factor(df.plot$veg_age);
str(df.plot)
df.plot$Element<-str_remove_all(df.plot$Element, ".pXRF") #remove .pXRF from variable names

#create table of means and se for graphing
mean.metals <- aggregate(PPM ~ Element+veg_age+layer_bot, df.plot, function(x) c(mean = mean(x)), na.action=na.omit); mean.metals$Element<-as.factor(mean.metals$Element)
str(mean.metals)
se.metals<- aggregate(PPM ~ Element+veg_age+layer_bot, df.plot, function(x) c(se = sd(x)/sqrt(length(x))), na.action=na.omit);se.metals;
tab.metals<-cbind(mean.metals, se.metals$PPM);names(tab.metals)[2:5]<-c("Months", "Depth", "PPM", "SE")
str(tab.metals)
                                                                                              
#graph metals PPM
theme_set(theme_base(base_size = 12))

p.metals.age<-ggplot(data=tab.metals, aes(x=Months, y=PPM, fill=Depth)) +
    # geom_hline(yintercept=100, linetype=1, color="darkgray", size=1)+
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Depth),color="black", shape=21)+
 facet_wrap(Element~., scale="free", ncol=3)


#graph sample position, depth, months
mean.metals.pos <- aggregate(PPM ~ Element+veg_age+sample_position, df.plot, function(x) c(mean = mean(x)), na.action=na.omit); mean.metals$Element<-as.factor(mean.metals$Element)
str(mean.metals.pos)
se.metals.pos<- aggregate(PPM ~ Element+veg_age+sample_position, df.plot, function(x) c(se = sd(x)/sqrt(length(x))), na.action=na.omit);se.metals;
tab.metals.pos<-cbind(mean.metals.pos, se.metals.pos$PPM);names(tab.metals.pos)[2:5]<-c("Months",  "Position", "PPM", "SE")
str(tab.metals.pos)

df.As<-subset(tab.metals.pos, Element=="As")

p.As<-ggplot(data=df.As, aes(x=Months, y=PPM, fill=Position)) +
    geom_hline(yintercept=20, linetype=2, color="red", size=1)+ #contaminated
    #geom_hline(yintercept=100, linetype=1, color="lightgray", size=1)+ #range
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="As (ppm)")+
  ylim(0,22)


#Cr graph
df.Cr<-subset(tab.metals.pos, Element=="Cr")
#facet_wrap(Depth~., scale="free")
p.Cr<-ggplot(data=df.Cr, aes(x=Months, y=PPM, fill=Position)) +
    geom_hline(yintercept=100, linetype=2, color="red", size=1)+ #contaminated
    geom_hline(yintercept=70, linetype=1, color="lightgray", size=1)+ #range
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Cr (ppm)")+
  ylim(0,210)

#Cu graph
df.Cu<-subset(tab.metals.pos, Element=="Cu")
p.Cu<-ggplot(data=df.Cu, aes(x=Months, y=PPM, fill=Position)) +
    geom_hline(yintercept=100, linetype=2, color="red", size=1)+ #contaminated
    geom_hline(yintercept=30, linetype=1, color="lightgray", size=1)+ #range
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Cu (ppm)")+
  ylim(0,110)

#Ni graph
df.Ni<-subset(tab.metals.pos, Element=="Ni")
p.Ni<-ggplot(data=df.Ni, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=50, linetype=1, color="lightgray", size=1)+ #range
    geom_hline(yintercept=50, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Ni (ppm)")+
  ylim(0,60)

#Pb graph
df.Pb<-subset(tab.metals.pos, Element=="Pb")
p.Pb<-ggplot(data=df.Pb, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=35, linetype=1, color="lightgray", size=1)+ #range
    geom_hline(yintercept=100, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Pb (ppm)")+
  ylim(0,110)

#Mn graph
df.Mn<-subset(tab.metals.pos, Element=="Mn")
p.Mn<-ggplot(data=df.Mn, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=1000, linetype=1, color="lightgray", size=1)+ #range
    geom_hline(yintercept=2000, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Mn (ppm)")+
  ylim(0,2010)

#Zn graph
df.Zn<-subset(tab.metals.pos, Element=="Zn")
p.Zn<-ggplot(data=df.Zn, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=90, linetype=1, color="lightgray", size=1)+ #range
    geom_hline(yintercept=300, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Zn (ppm)")+
  ylim(0,310)

#Mg graph
df.Mg<-subset(tab.metals.pos, Element=="Mg")
p.Mg<-ggplot(data=df.Ni, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=5000, linetype=1, color="lightgray", size=1)+ #range
   # geom_hline(yintercept=50, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Mg (ppm)")+
  ylim(0,5000)

#Fe graph
df.Fe<-subset(tab.metals.pos, Element=="Fe")
p.Fe<-ggplot(data=df.Fe, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=40000, linetype=1, color="lightgray", size=1)+ #range
    geom_hline(yintercept=50000, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Fe (ppm)")+
  ylim(0,50100)

#Ca graph
df.Ca<-subset(tab.metals.pos, Element=="Ca")
p.Fe<-ggplot(data=df.Fe, aes(x=Months, y=PPM, fill=Position)) +
      geom_hline(yintercept=40000, linetype=1, color="lightgray", size=1)+ #range
    geom_hline(yintercept=50000, linetype=2, color="red", size=1)+ #contaminated
    geom_errorbar(aes(ymin=PPM-SE, ymax=PPM+SE), width=.1)+
  geom_point(size=3, aes(fill = Position),color="black", shape=21)+
  labs(y="Fe (ppm)")+
  ylim(0,50100)


```
  
```{r, extra code, eval=FALSE, echo=FALSE}  

   
#Attempted
### Attempted outlier remove and transformation of Pb and Zn - Zn after 2 high outliers removed and log transformed = normal
#remove NAs in Pb and Zn
y.Pb<-as.numeric(na.omit(df.vine$Pb.pXRF))

#Outlier Test
chisq.out.test(y.Pb, variance=var(y.Pb), opposit =TRUE) #true = min value false=max
chisq.out.test(y.Pb, variance=var(y.Pb), opposit =FALSE) #higest value remove first as this is more significant, then retest normality
y.Pb[y.Pb > 29] <- NA
y.Pb[y.Pb < 13.00001] <- NA
y.Pb
#p-value = <.0001 highest value 30 is an outlier
shapiro.test((y.Pb)^3)
#check for outliers remove
str(y.Pb)
hist(y.Pb)

#remove NAs in Pb and Zn
y.Zn<-as.numeric(na.omit(df.vine$Zn.pXRF))
#Outlier Test for Zn
chisq.out.test(y.Zn, variance=var(y.Zn), opposit =TRUE) #true = min value false=max
chisq.out.test(y.Zn, variance=var(y.Zn), opposit =FALSE) #higest value remove first as this is more significant, then retest normality
y.Zn[y.Zn > 150] <- NA
#p-value = <.0001 highest value 30 is an outlier
shapiro.test(log(y.Zn))
#log(Zn) or log(Zn+1) are normal

#### box-cox tranformation of all metals

#create a function for boxcox transformation
powerTransform <- function(y, lambda1, lambda2 = NULL, method = "boxcox") {
  boxcoxTrans <- function(x, lam1, lam2 = NULL) {
    # if we set lambda2 to zero, it becomes the one parameter transformation
    lam2 <- ifelse(is.null(lam2), 0, lam2)
    if (lam1 == 0L) {
      log(y + lam2)
    } else {
      (((y + lam2)^lam1) - 1) / lam1 }}
  switch(method
         , boxcox = boxcoxTrans(y, lambda1, lambda2)
         , tukey = y^lambda1) }

#remove columns with all zeros, and Sn with only 1 value
colnames(df.vine)


#metals<-cbind(Al.pXRF, As.pXRF, Ca.pXRF, Co.pXRF, Cr.pXRF, Cu.pXRF, Fe.pXRF, Hg.pXRF, K.pXRF, LE.pXRF, Mg.pXRF, Mn.pXRF, Mo.pXRF, Nb.pXRF, Ni.pXRF, P.pXRF, Pb.pXRF, Rb.pXRF, S.pXRF,  Se.pXRF, Si.pXRF,  Sr.pXRF, Th.pXRF, Ti.pXRF, U.pXRF, V.pXRF, W.pXRF, Y.pXRF, Zn.pXRF, Zr.pXRF)

lm.m<-lm(m ~ veg_age*layer_bot*sample_position, data = df.vine)

bc<-boxcox(lm.m, na.omit=T) #run boxcox trans on linear model
lambda <- bc$x[which.max(bc$y)]

#re-run with transformation
   #mod.act3.bc <- lm(powerTransform(m.y, lambda) ~ Enzyme*Matrix*pH, data=df.m,
                  #na.action = na.omit)

#manova model -DATA NOT NORMAL YET
  #lm.metals1<-lm(y ~ layer_bot*sample_position*veg_age, data=df.m)
#manova model with vineyard as a random effect (block)
  #lmer.metals<-lmer (as.matrix(df.m[4:12]) ~ layer_bot*sample_position + (1|veg_age), data=df.m)
  #lmer.nest<-lmer (cbind(df.vine$ ~ layer_bot*sample_position + (1|veg_age), data=df.vine)
  #anova(lm.metals1)
  #summary(aov(lm.metals1))

#pairwise comparisions of means
cont1<-lsmeans(Permanova.metals, pairwise ~ layer_bot);
cont1.table<-summary(cont1$contrasts); cont1
names(m2) 

```

